---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r, echo=TRUE}
library(ggplot2)
```

## Loading and preprocessing the data

First, we'll download the data and unzip in our working directory

```{r, echo=TRUE}
fileURL <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
fileName <- "activity.zip"

#download the file:
download.file(fileURL, destfile = fileName, method = "curl")

#unzip the file:
unzip(fileName)
```

Next, we'll read the data into R and prep it for analysis
```{r, echo=TRUE}
#read the csv
activity_data <- read.csv("activity.csv", na.strings = "NA")

#convert date to date class
activity_data$date <- as.Date(activity_data$date, "%Y-%m-%d")

#convert interval to a factor
activity_data$interval <- as.factor(activity_data$interval)
```


## What is the average daily activity pattern?

Here we're taking a look at the distribution of total number of steps per day. Note that missing values (NA's) are being removed.

```{r, echo=TRUE}
#use tapply to calculate the total per day
total_steps <- tapply(activity_data$steps, activity_data$date, sum,  na.rm = TRUE)

#plot histogram showing the distribution of total steps per day
qplot(total_steps, binwidth=1000, xlab="Total Number of Steps", ylab = "Frequency", main = "Distribution of Total Steps Per Day")
```

Next, let's look at the mean & median of the total steps per day.

```{r, echo=TRUE}
#calculate the mean
mean(total_steps, na.rm = TRUE)
```
The median is 9,354.

```{r, echo=TRUE}
#calculate the median
median(total_steps, na.rm = TRUE)
```
The median is 10,395.

is.nan(activity_data$steps)

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?